\documentclass[12pt]{article}


\newcommand{\s}[1]{\texttt{#1}}
\renewcommand{\it}[1]{\textit{#1}}
\begin{document}

\title{Plan for Language Specification}
\author{Dhruv Rajan}
\date{\today}
\maketitle

\section{Figaro's Representation}
\subsection{Primitive Types}
In Figaro, every data structure is an \it{element}. Every element
holds a \textit{value}, with some associated \textit{value type}. Thus, all
elements are of the form \s{Element[type]}. Distinctions are made between four categories of elements:

\begin{enumerate}
\item \it{atomic} self contained, does not depend on another element (\s{Normal})
\item \it{compound} build out of other elements (\s{If (...), Apply (...)})
\item \it{discrete} element whose value type is discrete (\s{Poisson})
\item \it{continuous} element whose value type is continuous (\s{Gamma})
\end{enumerate}

\subsection{Observing Variables}
Elements are defined by distributions of their possible values. A
simple \s{Normal} element can only hold values on the interval
$[-1, 1]$, with corresponding probabilities. Thus, each element may be
considered a random variable, on its given distribution. Figaro
provides methodology for symbolic manipulation of these random
variables, allowing dependent variables (compound elements) to be
utilized. \\

No element is given an immediate value until it is observed. The
\s{observe()} method alters only the element on which it is called, so
that when an inference algorithm (such as
\s{VariableElimination.probability}) is run on that element, or any
variable in the same dependency network, it can calculate
the relevant conditional probabilities.

\subsection{Creating Compound Elements}
Atomic elements can be combined to form compound elements. The
simplest method for combining them is the \s{If} construct, which
allows for slightly complex conditioning on distributions

\subsubsection{If}
The type signature for \s{If} is as follows:

\begin{center}
  \s{If (test: Element[Boolean], \\ then: Element[T], \\ else:
    Element[T])\\ => cond: Element[T]}
\end{center}

This constructor creates an element which follows the corresponding distribution.

\subsubsection{Apply}
The type signature for \s{Apply} is as follows:
\begin{center}
  \s{Apply (e1: Element[T], \\ fn: T -> U)}
\end{center}

\subsubsection{Chain}
The type signature for \s{Chain} is as follows:
\begin{center}
  \s{Chain ((e1: Element[T], \\ fn: T -> Element[U]))}
\end{center}

\section{Haskell Implementation (src/Probability.hs)}
I've tried to take the core ideas in from the Figaro language, and express
them in Haskell. This includes, (1) Representation of Elements / Random Variables,
(2) Ability to observe Random Variables (3) Ability to use inference to infer
probability distributions, given observations.

\subsection{Representing Distributions}
I've tried to separate the complexity of combining random variables from
the representation of distributions. Thus, a distribution is just a function,
mapping some values of some type $a$ to probabilities (\s{Double}). I have yet
to integrate the standard library of distributions from the statistics library.
\begin{center}
  \s{type Distribution a = a -> Double}
\end{center}

\subsection{Representing Atomic Elements}
There is an Element type for representing the structure which correspond
to Figaro elements. Atomic elements have two attributes---one for the distribution
on which their values are drawn, and one to keep track of the element's observed
values.
\begin{center}
  \s{data Element a = Atomic \{distribution :: Distribution a, observed :: Maybe a\}}
\end{center}

Thus, observations can be performed on atomic elements simply by
changing this attribute.

\subsection{Creating Compound Elements}
The \s{Element} datatype has separate recursive type constructors for creating
compound elements from other elements. \s{If}, for example, creates an Element
from three existing elements. \s{Apply} and \s{Chain} have yet to be implemented.

\subsection{Inference}
The \s{probability} function contains simple inference logic. Since
compound elements are allowed, every element can be considered part of
a dependency tree. This function crawls down the tree recursively, calculating
the probabilities at each required step. The next steps will include looking
into more sophisticated inference algorithms.
\end{document}